StateSpaceModels.jl: a Julia Package for Time-Series Analysis in a State-Space Framework

### 2019-10-18 paper/paper.tex
Correct typo
updating paper

The remainder of this paper is organized as follows. In Section \ref{sec:statespace}, the Gaussian state-space framework is introduced. In Section \ref{sec:modeling}, the procedures for model specification are presented. Section \ref{sec:estimation} explains the filtering, estimation and smoothing processes. In Section \ref{sec:forecasting}, the forecasting and simulation procedures are presented. Section \ref{sec:applications} contains examples of applications that illustrate the use of the package. Finally, conclusions and future work are provided in Section \ref{sec:conclusion}

@@

Apart from the well-known predefined models presented in Section \ref{subsec:predefined}, the package allows the input of any user-defined linear model through the definition of matrices $Z_{t}$, $T$, and $R$ through the following constructor:
%
\begin{lstlisting}[language = Julia]
model = StateSpaceModel(y, Z, T, R)
\end{lstlisting}

In the simpler case where $Z_{t}$ is constant, the argument \texttt{Z} can be input as an \texttt{Array\{Float64, 2\}}. Conversely, if $Z_{t}$ is time-varying, it must be input as an \texttt{Array\{Float64, 3\}} where the third dimension represents the time periods.

@@

The choice of the filter type must be done in the estimation step as follows:
%
\begin{lstlisting}[language = Julia]
ss = statespace(model; filter_type = SquareRootFilter)
\end{lstlisting}

If no \texttt{filter\_type} is provided, the standard type \texttt{KalmanFilter} is utilized as a default option. Further information on Kalman filtering can be found in Chapter 4.3 of \cite{durbin2012time}.

@@

@@

In this case, our objective is to simulate a large set of future scenarios that can be used as inputs in a stochastic optimization problem with the goal of obtaining the best contracting strategy for a distribution company. To this end, the function \texttt{simulate} will be used to perform Monte Carlo simulation. We will also use the square-root Kalman filter for this application.
%
\begin{lstlisting}[language = Julia]
# Specify the state-space model
model = structural(consumption, 12; X = temperature)
# Estimate the state-space model
ss = statespace(model; filter_type = SquareRootFilter)
# Number of months ahead to be simulated
N = 24
# Number of scenarios to be simulated
S = 1000
# Perform simulation
sim = simulate(ss, N, S)
\end{lstlisting}

@@

\begin{lstlisting}[language = Julia]
# State transition matrix
T = kron(Matrix{Float64}(I, p, p), [1 (1 - ρ * Δ / 2) * Δ; 0 (1 - ρ * Δ)])
# Input matrix
R = kron(Matrix{Float64}(I, p, p), [.5 * Δ^2; Δ])
# Output (measurement) matrix
Z = kron(Matrix{Float64}(I, p, p), [1 0])
# User defined model
model = StateSpaceModel(y, Z, T, R)
# Estimate vehicle speed and position
ss = statespace(model)
\end{lstlisting}

In this example, we define the noise variances $H$ and $Q$, generate the noises and simulate a random vehicle trajectory using the state-space equations:
%
\begin{lstlisting}[language = Julia]
# Generate random actuators
Q = .5 * Matrix{Float64}(I, q, q)
η = MvNormal(zeros(q), Q)
# Generate random measurement noise
H = 2. * Matrix{Float64}(I, p, p)
ε = MvNormal(zeros(p), H)
# Simulate vehicle trajectory
α = Matrix{Float64}(undef, n + 1, m)
y = Matrix{Float64}(undef, n, p)
for t in 1:n
    y[t, :] = Z * α[t, :] + rand(ε)
    α[t + 1, :] = T * α[t, :] + R * rand(η)  
end
\end{lstlisting}

>>>>

The remainder of this paper is organized as follows. In Section \ref{sec:statespace}, the Gaussian state-space framework is introduced. In Section \ref{sec:modeling}, the procedures for model specification are presented. Section \ref{sec:estimation} explains the filtering, estimation and smoothing processes. Section \ref{sec:diagnostics} discusses the available diagnostics to evaluate the model specification and estimation. In Section \ref{sec:forecasting}, the forecasting and simulation procedures are presented. Section \ref{sec:applications} contains examples of applications that illustrate the use of the package. Finally, conclusions and future work are provided in Section \ref{sec:conclusion}.

@@

Apart from the well-known predefined models presented in Section \ref{subsec:predefined}, the package allows the input of any user-defined linear model through the definition of matrices $Z_{t}$, $T$, and $R$ through the following constructor:
%
\begin{lstlisting}[language = Julia]
model = StateSpaceModel{Float64}(y, Z, T, R)
\end{lstlisting}

In the simpler case where $Z_{t}$ is constant, the argument \texttt{Z} can be input as an \texttt{Array\{Float64, 2\}}. Conversely, if $Z_{t}$ is time-varying, it must be input as an \texttt{Array\{Float64, 3\}} where the third dimension represents the time periods. Any desired float precision can be employed within the model.

@@

The choice of the filter type must be done in the estimation step as follows:
%
\begin{lstlisting}[language = Julia]
ss = statespace(model; filter_type = SquareRootFilter{Float64})
\end{lstlisting}

Filters allow any desired float precision. If no \texttt{filter\_type} is provided, the standard type \texttt{KalmanFilter\{Float64\}} is utilized as a default option. Further information on Kalman filtering can be found in Chapter 4.3 of \cite{durbin2012time}.

@@

\section{Diagnostics} \label{sec:diagnostics}

As a way of evaluating the model specification and estimation, the package contains statistical diagnostics that can be conducted over the residuals. The available diagnostics are the Jarque-Bera normality test \cite{jarque1980efficient}, the Ljung-Box independence test \cite{box1970distribution}, and an homoscedasticity test \cite{durbin2012time}. The diagnostics can be run through function \texttt{diagnostics}.

An example of the diagnostics log can be seen below.
%
\begin{lstlisting}[language = Julia]
julia> diagnostics(ss)
==============================================================
                    Running diagnostics...
                Jarque-Bera: p-value = 0.89683
                  Ljung-Box: p-value = 0.95855
           Homoscedasticity: p-value = 0.30752
==============================================================
\end{lstlisting}

@@

In this case, our objective is to simulate a large set of future scenarios that can be used as inputs in a stochastic optimization problem with the goal of obtaining the best contracting strategy for a distribution company. To this end, the function \texttt{simulate} will be used to perform Monte Carlo simulation. We will also use the square-root Kalman filter for this application.
%
\begin{lstlisting}[language = Julia]
# Specify the state-space model
model = structural(consumption, 12; X = temperature)
# Estimate the state-space model
ss = statespace(model; filter_type = SquareRootFilter{Float64})
# Number of months ahead to be simulated
N = 24
# Number of scenarios to be simulated
S = 1000
# Perform simulation
sim = simulate(ss, N, S)
\end{lstlisting}

@@

\begin{lstlisting}[language = Julia]
# State transition matrix
T = kron(Matrix{Float64}(I, p, p), [1 (1 - ρ * Δ / 2) * Δ; 0 (1 - ρ * Δ)])
# Input matrix
R = kron(Matrix{Float64}(I, p, p), [.5 * Δ^2; Δ])
# Output (measurement) matrix
Z = kron(Matrix{Float64}(I, p, p), [1 0])
# User defined model
model = StateSpaceModel{Float64}(y, Z, T, R)
# Estimate vehicle speed and position
ss = statespace(model)
\end{lstlisting}

In this example, we define the noise variances $H$ and $Q$, generate the noises and simulate a random vehicle trajectory using the state-space equations.
%
\begin{lstlisting}[language = Julia]
# Generate random actuators
Q = .5 * Matrix{Float64}(I, q, q)
η = MvNormal(zeros(q), Q)
# Generate random measurement noise
H = 2. * Matrix{Float64}(I, p, p)
ε = MvNormal(zeros(p), H)
# Simulate vehicle trajectory
α = Matrix{Float64}(undef, n + 1, m)
y = Matrix{Float64}(undef, n, p)
for t in 1:n
    y[t, :] = Z * α[t, :] + rand(ε)
    α[t + 1, :] = T * α[t, :] + R * rand(η)  
end
\end{lstlisting}
